---
title: "Filtering Parameter Playground"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, message=FALSE, warning=FALSE}
find_package_root <- function(start = getwd()) {
  path <- normalizePath(start, mustWork = TRUE)
  while (TRUE) {
    if (file.exists(file.path(path, "DESCRIPTION")) && dir.exists(file.path(path, "R"))) {
      return(path)
    }
    parent <- dirname(path)
    if (identical(parent, path)) {
      stop("Could not find package root containing DESCRIPTION and R/.")
    }
    path <- parent
  }

  return(path)
}

package_root <- find_package_root()
knitr::opts_knit$set(root.dir = package_root)
devtools::load_all(package_root, quiet = TRUE)
```

This playground uses the package test MOO object and runs the public
`filter_counts()` wrapper. The examples below keep the CPM filtering settings the
same and vary only the total sample minimum from 8 to 9.

## Load Test Data

```{r load-test-data}
moo <- readr::read_rds("tests/testthat/data/moo.rds")

count_type <- "clean"
feature_id_colname <- "GeneName"
sample_id_colname <- "Sample"
group_colname <- "Group"
label_colname <- "Label"

sample_metadata <- as.data.frame(moo@sample_meta)
counts_before <- as.data.frame(moo@counts[[count_type]])
sample_columns <- setdiff(colnames(counts_before), feature_id_colname)

knitr::kable(data.frame(
  count_type = count_type,
  features_before_filtering = nrow(counts_before),
  samples = length(sample_columns),
  feature_id_colname = feature_id_colname,
  sample_id_colname = sample_id_colname,
  group_colname = group_colname,
  check.names = FALSE
))
```

## Filtering Parameters

```{r filtering-parameters}
minimum_count_value_to_be_considered_nonzero <- 4
minimum_number_of_samples_with_nonzero_counts_in_a_group <- 2
use_cpm_counts_to_filter <- TRUE
use_group_based_filtering <- FALSE
```

## CPM Passing-Sample Summary

This table counts how many samples pass the CPM threshold for each feature. A
feature kept with a total minimum of 8 can still have low values in up to two
samples.

```{r cpm-passing-summary}
cpm_counts <- counts_before
cpm_counts[, sample_columns] <- edgeR::cpm(as.matrix(counts_before[, sample_columns]))

passing_sample_counts <- rowSums(
  as.matrix(cpm_counts[, sample_columns]) >= minimum_count_value_to_be_considered_nonzero
)

knitr::kable(data.frame(
  passing_samples = sort(unique(passing_sample_counts)),
  features = as.integer(tabulate(
    match(passing_sample_counts, sort(unique(passing_sample_counts))),
    nbins = length(unique(passing_sample_counts))
  )),
  kept_when_total_minimum_is_8 = sort(unique(passing_sample_counts)) >= 8,
  kept_when_total_minimum_is_9 = sort(unique(passing_sample_counts)) >= 9,
  check.names = FALSE
))
```

## Run `filter_counts()`: Total Minimum 8

```{r run-filter-counts-threshold-5-total-5, message=TRUE}
filtered_moo_total_8 <- filter_counts(
  moo,
  count_type = count_type,
  feature_id_colname = feature_id_colname,
  sample_id_colname = sample_id_colname,
  group_colname = group_colname,
  label_colname = label_colname,
  minimum_count_value_to_be_considered_nonzero = 4,
  minimum_number_of_samples_with_nonzero_counts_in_total = 8,
  minimum_number_of_samples_with_nonzero_counts_in_a_group = 2,
  use_cpm_counts_to_filter = use_cpm_counts_to_filter,
  use_group_based_filtering = use_group_based_filtering,
  print_plots = TRUE,
  save_plots = FALSE
)
```

```{r run-filter-counts-total-8, message=TRUE}
filtered_moo_total_8 <- filter_counts(
  moo,
  count_type = count_type,
  feature_id_colname = feature_id_colname,
  sample_id_colname = sample_id_colname,
  group_colname = group_colname,
  label_colname = label_colname,
  minimum_count_value_to_be_considered_nonzero = 5,
  minimum_number_of_samples_with_nonzero_counts_in_total = 5,
  minimum_number_of_samples_with_nonzero_counts_in_a_group = 2,
  use_cpm_counts_to_filter = use_cpm_counts_to_filter,
  use_group_based_filtering = use_group_based_filtering,
  print_plots = TRUE,
  save_plots = FALSE
)
```


## Run `filter_counts()`: Total Minimum 9

```{r run-filter-counts-total-9, message=TRUE}
filtered_moo_total_9 <- filter_counts(
  moo,
  count_type = count_type,
  feature_id_colname = feature_id_colname,
  sample_id_colname = sample_id_colname,
  group_colname = group_colname,
  label_colname = label_colname,
  minimum_count_value_to_be_considered_nonzero = 4,
  minimum_number_of_samples_with_nonzero_counts_in_total = 9,
  minimum_number_of_samples_with_nonzero_counts_in_a_group = 2,
  use_cpm_counts_to_filter = use_cpm_counts_to_filter,
  use_group_based_filtering = use_group_based_filtering,
  print_plots = TRUE,
  save_plots = FALSE
)
```


```{r run-interactive, message=TRUE}
filtered_moo_total_9 <- filter_counts(
  moo,
  count_type = count_type,
  feature_id_colname = feature_id_colname,
  sample_id_colname = sample_id_colname,
  group_colname = group_colname,
  label_colname = label_colname,
  minimum_count_value_to_be_considered_nonzero = 4,
  minimum_number_of_samples_with_nonzero_counts_in_total = 9,
  minimum_number_of_samples_with_nonzero_counts_in_a_group = 2,
  use_cpm_counts_to_filter = use_cpm_counts_to_filter,
  use_group_based_filtering = use_group_based_filtering,
  interactive_plots = T,
  print_plots = TRUE,
  save_plots = FALSE
)
```

## 8 vs 9 Summary

```{r compare-total-8-total-9}
counts_total_8 <- as.data.frame(filtered_moo_total_8@counts[["filt"]])
counts_total_9 <- as.data.frame(filtered_moo_total_9@counts[["filt"]])

features_total_8 <- counts_total_8[[feature_id_colname]]
features_total_9 <- counts_total_9[[feature_id_colname]]
features_kept_only_by_8 <- setdiff(features_total_8, features_total_9)

knitr::kable(data.frame(
  setting = c("total minimum 8", "total minimum 9", "kept by 8 only"),
  features = c(
    length(features_total_8),
    length(features_total_9),
    length(features_kept_only_by_8)
  ),
  check.names = FALSE
))
```

```{r low-values-kept-only-by-8}
counts_kept_only_by_8 <- counts_total_8 |>
  dplyr::filter(.data[[feature_id_colname]] %in% features_kept_only_by_8)

if (nrow(counts_kept_only_by_8) > 0) {
  low_value_counts <- rowSums(
    as.matrix(counts_kept_only_by_8[, sample_columns]) < minimum_count_value_to_be_considered_nonzero
  )

  knitr::kable(data.frame(
    low_values_per_feature = sort(unique(low_value_counts)),
    features_kept_only_by_8 = as.integer(tabulate(
      match(low_value_counts, sort(unique(low_value_counts))),
      nbins = length(unique(low_value_counts))
    )),
    check.names = FALSE
  ))
} else {
  knitr::kable(data.frame(message = "No features were kept by 8 and removed by 9."))
}
```
